### 2013-07-13 doc/sosp13/evaluation.tex
incorporate todos reviews

The goal of our evaluation is to show that the core pieces of the \Grappa
runtime system, namely our tasking system and the global memory/communication
layer work as expected and together are able to efficiently run irregular
applications. We evaluate \Grappa in three basic steps:

@@

\item We present results that show that the \Grappa runtime is able to sustain
very high concurrency rates and the communication layer is able to sustain a
very high rate of global memory operations. We also show the performance of a
graph kernel that stresses communication and concurrency together.

\item We show how some popular irregular applications running on \Grappa
compare to the Cray XMT and hand-tuned MPI code.

\item We finish with a characterization of system behavior, including
profiling where execution time goes, how aggregation affects message size and
rates, how global memory and work stealing behaves.

@@

\paragraph{Putting it all together with Unbalanced Tree Search in memory
(UTS).} 
Figure~\ref{fig:grappa-uts} shows the overall performance of \Grappa running
UTS. The point of this experiment is to demonstrate whether \Grappa's context
switching and communication layers can in fact be used together, while
balancing workload, to run an irregular application efficiently. Visiting
vertices in the distributed tree requires mostly remote accesses, and because
each vertex in the tree must be visited before it can be expanded, blocking
remote reads are required. 
In this case, we are forced to context switch to tolerate the remote access and continue aggregating. 
We look at two classes of trees, T1 and T3, from
the original benchmark. T1 trees are very shallow and wide (i.e., significant
parallel slack), while T3 trees are very deep (i.e., little parallel slack).
When the access to each vertex is a random access, the critical path to search
T3 trees is very long, hence the low performance and scalability. On such
trees, we do not expect there to be sufficient concurrency for any system,
including \Grappa, to achieve high throughput --- at the 16-node data point,
the average active tasks per core over the search is 775 and 13 for T1 and T3,
respectively. On the other hand \Grappa performs and scales very well for T1
trees.

@@

@@

\begin{figure}[ht]
    \begin{center}
      \begin{tabular}{c|c c c c}
        Benchmark     & Comm & User & Idle & Sched \\ \hline
        GUPS          & 62.9  & 18.9    & 15.4 & 2.80 \\ 
        BFS           & 54.84 & 30.90   & 10.94 & 3.43 \\ 
        Intsort       & 34.28 & 42.00   & 21.31 & 2.47 \\ 
        UTS           & 40.57 & 56.52   &  1.21 & 1.73 \\
        Pagerank      & 76.79 & 20.71   &  0.06 & 2.50 \\
      \end{tabular}
    \end{center}
    \caption{\Grappa\ execution time profile, in percent.}
%bar chart, one set of bars per benchmark, one bar per system. runtime normalized to grappa.
    \label{fig:grappa-profile}
\end{figure}

@@

In BFS, we see a trimodal distribution. There are two reasons for this. First, in BFS, we obtained the best throughput by flushing at least 500 microseconds apart, and also flushing whenever a core is idle. Since this process ignores whether another flush has occurred recently, it leads to a large number of small messages. This leads to the first and smallest mode of the distribution.  The other two modes are due to a locality optimization in the \Grappa version of BFS. To minimize unnecessary remote communication, each core visits vertices whoes edge lists are stored locally. Since we are traversing power-law graphs, a small fraction of vertices have a very high number of edges. This means that occasionally a core visits a high-degree vertex which allows it to very quickly generate a large batch of requests. This interacts well with aggregation; when we find high degree vertices, we are able to aggregate packets on the order of 40KB. In flush intervals without high-degree vertices, we see packets on the order of 8KB.

>>>>

Our questions to answer with experimentation are:
The goal of our evaluation is to show that the core pieces of the \Grappa
runtime system, namely our tasking system and the global memory/communication
layer work as expected and together are able to efficiently run irregular
applications. We evaluate \Grappa in three basic steps:

@@

\item We present results that show that \Grappa can
    support large amounts of concurrency: sufficient for remote memory
    access and aggregation. The communication layer is able to
    achievesustain a very high rate of global memory operations. We also show the performance of a
graph kernel that stresses communication and concurrency together.
\TODO{compare to key-value stores?}

\item We characterize system behavior, including
profiling where execution time goes, how aggregation affects message size and
rates, and \checkme{the relationship between aggregation and
    concurrency.}

\item Finally, we show how some more realistic irregular workloads on \Grappa
compare to the Cray XMT and hand-tuned MPI code.

\TODO{Grappa is well-suited to applications where there is
    abundant data parallelism but unpredictable or imbalanced data
    access. BSP is efficient where there is sufficient parallel slack.
    Although Grappa loses some asynchrony by relying heavily on
    aggregation, it still provides more flexibility for sections of the
    computation to proceed more quickly.

\TODO{how much concurrency is needed on apps. Our basic plot of
    performance vs workers (with and without aggro?)}
\TODO{relationship between user-thread-visible latency and aggregate throughput.}
\TODO{characterization of aggregation message size vs nodes}
\TODO{show our scalability plots; if they don't scale, show
    specifically why. It is okay if its because aggregation stops
    working after a point if we can discuss future scalability.
    Perhaps reference ``MPI on millions of cores'' type work}

@@

\TODO{go into more detail about what Grappa is doing when executing
    GUPs. Specifically note what the rate per core is, what the
    aggregation and deaggregation rates are. Show on the plot the
    message size distribution (some box plot)}

\paragraph{Putting it all together with Unbalanced Tree Search in memory
(UTS).} 
Figure~\ref{fig:grappa-uts} shows the overall performance of \Grappa running
UTS. The point of this experiment is to demonstrate whether \Grappa's context
switching and communication layers can in fact be used together, while
balancing workload, to run an irregular application efficiently. Visiting
vertices in the distributed tree requires mostly remote accesses, and because
each vertex in the tree must be visited before it can be expanded, blocking
remote reads are required. 
In this case, we are forced to context switch to tolerate the remote access and continue aggregating. 
We look at two classes of trees, T1 and T3, from
the original benchmark. T1 trees are very shallow and wide (i.e., significant
parallel slack \TODO{cite Valiant and say more somewhere about BSP}), while T3 trees are very deep (i.e., little parallel slack).
When the access to each vertex is a random access, the critical path to search
T3 trees is very long, hence the low performance and scalability. On such
trees, we do not expect there to be sufficient concurrency for any system,
including \Grappa, to achieve high throughput --- at the 16-node data point,
the average active tasks per core over the search is 775 and 13 for T1 and T3,
respectively. On the other hand \Grappa performs and scales very well for T1
trees.

@@

\TODO{addressing reviwer2: sweep of irregularness in one aggregate or multiple dimensions
    and how a specific aspect of performance changes in different
    systems}

@@

\begin{figure}[ht]
    \begin{center}
      \begin{tabular}{c|c c c c}
        Benchmark     & Comm & User & Idle & Sched \\ \hline
        GUPS          & 62.9  & 18.9    & 15.4 & 2.80 \\ 
        BFS           & 54.84 & 30.90   & 10.94 & 3.43 \\ 
        Intsort       & 34.28 & 42.00   & 21.31 & 2.47 \\ 
        UTS           & 40.57 & 56.52   &  1.21 & 1.73 \\
        Pagerank      & 76.79 & 20.71   &  0.06 & 2.50 \\
      \end{tabular}
    \end{center}
    \caption{\Grappa\ execution time profile, in percent.}
%bar chart, one set of bars per benchmark, one bar per system. runtime normalized to grappa.
    \label{fig:grappa-profile}
\end{figure}
\TODO{ shorten the execution profile section in favor of deeper
    discussion about where the message rate comes from}

@@

In BFS, we see a trimodal distribution. There are two reasons for
this. First, in BFS, we obtained the best throughput by flushing at
least 500 microseconds apart, and also flushing whenever a core is
idle. Since this process ignores whether another flush has occurred
recently, it leads to a large number of small messages. This leads to
the first and smallest mode of the distribution.  The other two modes
are due to a locality optimization in the \Grappa version of BFS. To
minimize unnecessary remote communication, each core visits vertices
whose edge lists are stored locally. Since we are traversing power-law graphs, a small fraction of vertices have a very high number of edges. This means that occasionally a core visits a high-degree vertex which allows it to very quickly generate a large batch of requests. This interacts well with aggregation; when we find high degree vertices, we are able to aggregate packets on the order of 40KB. In flush intervals without high-degree vertices, we see packets on the order of 8KB.

### 2013-07-13 doc/sosp13/memory.tex
incorporate todos reviews

@@

>>>>

\TODO{eliminate the word cache from the paper since the main point is
    contiguity and not unpredictable temporal reuse. We
    must make it clear that this is not user-level caching.}

@@

\TODO{admit the lack of a synchronization feature for lightweight
    transactions across two or more domains. Talk about using the same
    philosophy of moving data along with the sync (same as a real
    cache) and utilizing latency tolerance. Then say further is out of
    scope for this paper.}
